---
title: "GRASSWORKS Project: Vegetation diversity analysis"
subtitle: "Question 2: Restoration factors: Response variable: FCSi (Index of Favourable Conservation Status)"
author: "<b>Christin Juno Laschke</b>"
date: today
date-format: iso
format:
  html:
    embed-resources: true
toc: true
toc-depth: 3
toc-expand: 2
number-sections: true
dev: png
fig-width: 7
fig-height: 5
---

# Preparation

## Load packages

```{r message=FALSE}
library(tidyverse)
library(here)
library(lattice)
library(performance) # visual check of model assumptions
library(emmeans) # calculate estimated marginal means and post-hoc Tukey
library(rstatix)
library(glmmTMB)
library(ggbeeswarm)
library(mgcv)
```

```{r include=FALSE}
rm(list = ls())
```

## Load data

```{r}
sites <- read_csv(
  here("data", "processed", "data_processed.csv"),
  col_names = TRUE, na = c("na", "NA", ""), col_types = cols(
    .default = "?"
  )) %>%
  dplyr::select(
    id.site, rest.meth, rest.age,
    ) %>%
  mutate(
    region = fct_relevel(region, "north", "centre", "south"),
    hydrology = fct_relevel(hydrology, "dry", "fresh", "moist"),
    rest.meth = fct_relevel(rest.meth, "cus", "mga", "res", "dih")
  )
```

# Data exploration

## a Missing values
```{r}
colSums(is.na(data_all))
```

66 NAs in rest.meth, rest.age and land.use.hist due to reference sites and 13 other NAs in rest.age


## Outliers

```{r}
Z <- data_all %>% 
  dplyr::select(where(is.numeric))
dotplot(as.matrix(Z), groups = FALSE,
        strip = strip.custom(bg = 'white',
                             par.strip.text = list(cex = 0.8)),
        scales = list(x = list(relation = "free"),
                      y = list(relation = "free"),
                      draw = FALSE),
        col = 1, cex  = 0.5, pch = 16,
        xlab = "Value of the variable",
        ylab = "Order of the data from text file")
data_all %>% 
  select(id.site, fcsi.hill.0) %>% 
  identify_outliers(fcsi.hill.0)
data_all %>% 
  select(id.site, rest.age.std) %>% 
  identify_outliers(rest.age.std)
```

## Check normality assumption


### Build linear model

```{r}
model  <- lm(rest.age ~ rest.meth, data = sites, na.action = na.omit)
```

### Create a QQ plot of residuals

```{r}
ggqqplot(residuals(model))
```

### Compute Shapiro-Wilk test of normality

```{r}
shapiro_test(residuals(model))
```

In the QQ plot, as all the points fall approximately along the reference line, we can assume normality. This conclusion is supported by the Shapiro-Wilk test. The p-value is not significant (p = 0.191), so we can assume normality.

### Check normality assumption by groups

```{r}
sites %>%
  filter(!is.na(rest.meth)) %>% 
  group_by(rest.meth) %>%
  shapiro_test(rest.age)
```

Note that, if your sample size is greater than 50, the normal QQ plot is preferred because at larger sample sizes the Shapiro-Wilk test becomes very sensitive even to a minor deviation from normality:

```{r}
ggqqplot(sites, "rest.age", facet.by = "rest.meth")
```
--> looks okay



## Check homogeneity of variance assumption


#### Residuals versus fits plot
```{r}
plot(model, 1)
```
# unclear if homogeneity is true

#### Levene's test

```{r}
sites %>% levene_test(rest.age ~ rest.meth)
```

From the output above, we can see that the p-value is < 0.05, which is significant. This means that, there is significant difference between variances across groups. Therefore, we can not assume the homogeneity of variances in the different treatment groups.



# Computation

```{r}
res_aov <- anova_test(model)
```

Include only sites with an age of less than 20 years

```{r}
sites_20y <- sites %>% 
  filter(rest.age <= 20)
model_20y  <- lm(rest.age ~ rest.meth, data = sites_20y, na.action = na.omit)
res_aov_20y <- anova_test(model_20y)
```

# Session info

```{r}
sessionInfo()
```